---
title: "Imputation(test)"
author: "Xinyi Lin"
date: "8/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The purpose of following codes is to test imputation ability of two packages--Himsc, missForest and mice.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(missForest)
library(Hmisc)
library(caret)
library(mice)
library(mi)
```

## Import data

Import "training.dta" as target dataset. 

```{r}
library(haven)
training <- read_dta("training.dta")
```

## Creat the variable

Notation:

Variable name: student_r

Created from: occup1_r, occup2_r

Label: student_r — yes-is a student, no-not a student, NA-occup1_r and occup2_r are NAs

```{r}
student_data = training %>%
  as_data_frame() %>%
  mutate(student_r = ifelse(occup1_r == 8 | occup2_r == 8, "yes", "no"),
         student_r = ifelse(is.na(student_r) & occup1_r == 20, "no", student_r))
```

```{r}
# for sexp1yr > 92, make them as NAs
impu_data = student_data %>% 
  filter(ageyrs_r <= 19) %>% 
  mutate(student_r = as.factor(student_r),
         visit = as.factor(visit),
         sex = ifelse(female==1, "female", "male")) %>% 
  select(ageyrs_r, sex, student_r, area, educate_r, currmarr_r, sexp1yr, SEScat) %>% 
  mutate(sex = as.factor(sex),
         area = as.factor(area),
         educate_r = as.factor(educate_r),
         currmarr_r = as.factor(currmarr_r),
         sexp1yr = ifelse(sexp1yr > 92, NA, sexp1yr),  
         SEScat = as.factor(SEScat))
head(impu_data)
```

# Test dataset

First, choose observations with known student_r as test dataset. There are 41482 observations in total.

```{r}
test_data = impu_data[!is.na(impu_data$student_r),]
```

Then, randomly split test dataset into ten subsets and set them as missing values to test error rates of each fold when using different methods.

```{r}
flds <- createFolds(1:41482, k = 10, list = TRUE, returnTrain = FALSE)
#flds
```

```{r, eval=FALSE}
set.seed(123)
test_data = as.data.frame(as.matrix(test_data))
na_data = test_data
na_data[flds[[1]], 3] = NA
mdf = missing_data.frame(na_data)
imputations <- mi(mdf)
```

```{r, results='hide'}
# error_rate = vector("list", 10)
set.seed(123)
test_data = as.data.frame(as.matrix(test_data))
error_himsc = rep(NA, 10)
error_forest = rep(NA, 10)
error_mice = rep(NA, 10)
for (n in 1:10){
  na_data = test_data
  na_data[flds[[n]], 3] = NA
  # missForest
  impu_forest = missForest(na_data)
  impu_forest_df = impu_forest$ximp
  # Himsc
  impu_himsc = aregImpute(~ ageyrs_r + sex + student_r + area + educate_r + currmarr_r + SEScat, data = na_data)
  impu_himsc_l = impute.transcan(impu_himsc, data=na_data, imputation=1, list.out=TRUE, pr=FALSE, check=FALSE)
  impu_himsc_df = as.data.frame(impu_himsc_l)
  # mice
  mice_data = mice(impu_data,seed = 123)
  impu_mice_df = complete(mice_data, 1)
  # calculate error rates
  # yes = 2, no = 1
  error_himsc[n] = sum(abs(as.numeric(impu_himsc_df[flds[[n]], 3]) - as.numeric(test_data[flds[[n]], 3])))/length(test_data[flds[[n]], 3])
  error_forest[n] = sum(abs(as.numeric(impu_forest_df[flds[[n]], 3]) - as.numeric(test_data[flds[[n]], 3])))/length(test_data[flds[[n]], 3])
  error_mice[n] = sum(abs(as.numeric(impu_mice_df[flds[[n]], 3]) - as.numeric(test_data[flds[[n]], 3])))/length(test_data[flds[[n]], 3])
}
res_error = list(himsc = error_himsc, forest = error_forest, mice = error_mice)
```

We can find that error rates of Himsc package is around 33% and error rates of missForest is around 24%. Following is the plot of error rates.

```{r}
res_error
res_error %>% 
  as.data.frame() %>% 
  mutate(fold = 1:10) %>% 
  gather(key = methods, value = error_rate, himsc:mice) %>% 
  ggplot(aes(x = fold, y = error_rate, color = methods)) + geom_line()
```

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```