---
title: "Imputation(add new variables)"
author: "Xinyi Lin"
date: "9/29/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(missForest)
library(ggplot2)
library(caret)
```

## Import data

Import data "R1_R18_partcleaned_20190913.dta".

```{r}
library(haven)
new_data <- read_dta("R1_R18_partcleaned_20190913.dta")
```

## Creat the variable

Notation:

Variable name: student_r

Created from: occup1_r, occup2_r

Label: student_r — yes-is a student, no-not a student, NA-occup1_r and occup2_r are NAs

```{r}
student_data = new_data %>%
  as.data.frame() %>% 
  mutate(student_r = ifelse(occup1_r == 8 | occup2_r == 8, "yes", "no"),
         student_r = ifelse(is.na(student_r) & occup1_r == 20, "no", student_r))
#head(student_data)
```

## Analysis missing data

### Select variables and tidy data

Predictors including: ageyrs_r, sex, student_r, area, educate_r, currmarr_r, sexp1yr, SEScat, study_id, pregnow_r, numchild_r, eversex_r, currrltn.

```{r}
target_data = student_data %>% 
  filter(ageyrs_r <= 19) %>% 
  mutate(student_r = as.factor(student_r),
         visit = as.factor(visit),
         sex = ifelse(female==1, "female", "male")) %>% 
  select(student_r, visit, ageyrs_r, sex, ageyrs_r, sex, student_r, area, educate_r, currmarr_r, sexp1yr, SEScat, study_id, pregnow_r, numchild_r, eversex_r, currrltn)
```

First, check each variable.

```{r}
summary(target_data)
```

For `sexp1yr`:

```{r}
summary(as.factor(target_data$sexp1yr))
```

93 means a lot, 97 means don't know, 98 means not applicable and 99 means not response. 97, 98, 99 can be marked as NA.

For `pregnow_r`:

```{r}
target_data %>% 
  group_by(sex, pregnow_r) %>% 
  dplyr::summarize(n = n())
```

For male, `pregnow_r` can be marked as 98. And `pregnow_r` should be catagorical variable.

For `currrltn`:

```{r}
summary(as.factor(target_data$currrltn))
```

7 means don't know. *0, 8, 9 can be marked as missing value(not sure).*

As there are too much missing value in `numchild_r`, and too much categories in `study_id` it is not suitable to be a predictor.

```{r}
impu_data = target_data %>% 
  mutate(sexp1yr = ifelse(sexp1yr %in% c(97, 98, 99), NA, sexp1yr)) %>% 
  mutate(pregnow_r = ifelse(sex=="male", 98, pregnow_r),
         pregnow_r = as.factor(pregnow_r)) %>% 
  mutate(currrltn = ifelse(currrltn %in% c(0, 8,9), NA, currrltn),
         currrltn = as.factor(currrltn)) %>% 
  select(-c(numchild_r, study_id)) %>% 
  mutate(visit = as.factor(visit),
         area = as.factor(area),
         educate_r = as.factor(educate_r),
         currmarr_r = as.factor(currmarr_r),
         SEScat = as.factor(SEScat),
         eversex_r = as.factor(eversex_r),
         sex = as.factor(sex))
summary(impu_data)
```

### Permutation 

Resample each variables separately to test their contribution to imputation.

```{r}
test_data = impu_data[!is.na(impu_data$student_r),]
```

```{r, results='hide'}
set.seed(123)
var_mis <- matrix(data=NA,nrow=10,ncol=10) # 10 variables and 10 folds for each variables
for (j in 3:12) {
  flds <- createFolds(1:nrow(test_data), k = 10, list = TRUE, returnTrain = FALSE)
  target_variable = test_data[,j]
  new_variable = sample(target_variable, nrow(test_data), replace = T)
  test_data[,j] = new_variable
  test_data = as.data.frame(as.matrix(test_data))
  for (n in 1:10) {
    na_data = test_data
    na_data[flds[[n]], 1] = NA
    # missForest
    impu_forest = missForest(na_data)
    impu_forest_df = impu_forest$ximp
    # calculate error rates
    # yes = 2, no = 1
    var_mis[j-2,n] = sum(abs(as.numeric(impu_forest_df[flds[[n]], 1]) - as.numeric(test_data[flds[[n]], 1])))/length(test_data[flds[[n]], 1])
  }
}
```

```{r}
colnames(var_mis) = c("fold1", "fold2", "fold3", "fold4", "fold5", "fold6", "fold7", "fold8", "fold9", "fold10")
rownames(var_mis) = c("ageyrs_r","sex","area","educate_r","currmarr_r","sexp1yr","SEScat","pregnow_r","eversex_r","currrltn")
apply(var_mis,1,mean) %>% 
  sort(decreasing = T)
```

### Test data

First, choose observations with known student_r as test dataset. There are `r nrow(test_data)` observations in total.

Then, randomly split test dataset into ten subsets and set them as missing values to test error rates of each fold when using different methods.

```{r}
flds <- createFolds(1:nrow(test_data), k = 10, list = TRUE, returnTrain = FALSE)
#flds
```

```{r, results='hide'}
# error_rate = vector("list", 10)
set.seed(123)
test_data = as.data.frame(as.matrix(test_data))
error_forest = rep(NA, 10) # indicate misclassification rate
error_index = rep(0,nrow(test_data)) # indcate misclassification rows
for (n in 1:10){
  na_data = test_data
  na_data[flds[[n]], 1] = NA
  # missForest
  impu_forest = missForest(na_data)
  impu_forest_df = impu_forest$ximp
  # calculate error rates
  # yes = 2, no = 1
  error_forest[n] = sum(abs(as.numeric(impu_forest_df[flds[[n]], 1]) - as.numeric(test_data[flds[[n]], 1])))/length(test_data[flds[[n]], 1])
  error_vector = abs(as.numeric(impu_forest_df[, 1]) - as.numeric(test_data[, 1]))
  error_index = error_index + error_vector
}
```

For misclassification rate:

```{r}
summary(error_forest)
plot(1:10,error_forest,xlab="Fold",ylab="Error Rate",type="b",lty=1)
```

For the misclassification dataset:

```{r}
error_data = test_data %>% 
  mutate(miscla = error_index,
         miscla = as.factor(miscla))
```

There are `r nrow(error_data)` observations are misclassified. Marginal distribution of misclassification rate in different gourps are significant among following variables: `educate_r`， `currmarr_r`, `sexp1yr`， `pregnow_r`.

For `educate_r`: **significant difference**

```{r}
educate_error = error_data %>% 
  group_by(miscla, educate_r) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(educate_r, mis_rate, sum) 
educate_error %>% knitr::kable()
educate_error %>% 
  ggplot(aes(educate_r, mis_rate)) +
  geom_bar(stat = "identity")
```

For `currmarr_r`: **significant difference**

```{r}
currmarr_error = error_data %>% 
  group_by(miscla, currmarr_r) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(miscla_1 = ifelse(is.na(miscla_1), 0, miscla_1)) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(currmarr_r, mis_rate, sum) 
currmarr_error %>% knitr::kable()
currmarr_error %>% 
  ggplot(aes(currmarr_r, mis_rate)) +
  geom_bar(stat = "identity")
```

For `sexp1yr`: **significant difference**

```{r}
sexp1yr_error = error_data %>% 
  group_by(miscla, sexp1yr) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(miscla_1 = ifelse(is.na(miscla_1), 0, miscla_1),
         miscla_0 = ifelse(is.na(miscla_0), 0, miscla_0)) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(sexp1yr, mis_rate, sum) 
sexp1yr_error %>% knitr::kable()
sexp1yr_error %>% 
  ggplot(aes(sexp1yr, mis_rate)) +
  geom_bar(stat = "identity")
```

For `pregnow_r`: **significant difference**

```{r}
pregnow_error = error_data %>% 
  group_by(miscla, pregnow_r) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(pregnow_r, mis_rate, sum) 
pregnow_error %>% knitr::kable()
pregnow_error %>% 
  ggplot(aes(pregnow_r, mis_rate)) +
  geom_bar(stat = "identity")
```

For `visit`:

```{r}
visit_error = error_data %>% 
  mutate(visit = as.numeric(visit)) %>% 
  group_by(miscla, visit) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(visit, mis_rate, sum) 
visit_error %>% knitr::kable()
visit_error %>% 
  ggplot(aes(visit, mis_rate)) +
  geom_bar(stat = "identity")
```

For `ageyrs_r`:

```{r}
age_error = error_data %>% 
  group_by(miscla, ageyrs_r) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(miscla_1 = ifelse(is.na(miscla_1), 0, miscla_1)) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(ageyrs_r, mis_rate, sum) 
age_error %>% knitr::kable()
age_error %>% 
  ggplot(aes(ageyrs_r, mis_rate)) +
  geom_bar(stat = "identity")
```

For `sex`:

```{r}
sex_error = error_data %>% 
  group_by(miscla, sex) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(sex, mis_rate, sum) 
sex_error %>% knitr::kable()
sex_error %>% 
  ggplot(aes(sex, mis_rate)) +
  geom_bar(stat = "identity")
```

For `area`:

```{r}
area_error = error_data %>% 
  group_by(miscla, area) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(area, mis_rate, sum) 
area_error %>% knitr::kable()
area_error %>% 
  ggplot(aes(area, mis_rate)) +
  geom_bar(stat = "identity")
```

For `SEScat`:

```{r}
SEScat_error = error_data %>% 
  group_by(miscla, SEScat) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(SEScat, mis_rate, sum) 
SEScat_error %>% knitr::kable()
SEScat_error %>% 
  ggplot(aes(SEScat, mis_rate)) +
  geom_bar(stat = "identity")
```

For `eversex`:

```{r}
eversex_error = error_data %>% 
  group_by(miscla, eversex_r) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(miscla_1 = ifelse(is.na(miscla_1), 0, miscla_1)) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(eversex_r, mis_rate, sum) 
eversex_error %>% knitr::kable()
eversex_error %>% 
  ggplot(aes(eversex_r, mis_rate)) +
  geom_bar(stat = "identity")
```

For `currrltn`:

```{r}
currrltn_error = error_data %>% 
  group_by(miscla, currrltn) %>% 
  dplyr::summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(miscla = str_c("miscla_", miscla)) %>% 
  spread(key = miscla, value = n) %>% 
  mutate(sum = miscla_0 + miscla_1,
         mis_rate = miscla_1/sum) %>% 
  select(currrltn, mis_rate, sum) 
currrltn_error %>% knitr::kable()
currrltn_error %>% 
  ggplot(aes(currrltn, mis_rate)) +
  geom_bar(stat = "identity")
```

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```