---
title: "Imputation"
author: "Xinyi Lin"
date: "7/30/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The goal of this file is to show the process and results of multiple imputation in the `student_r` variable.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(missForest)
library(Hmisc)
```

## Import data

Import "training.dta" as target dataset. 

```{r}
library(haven)
training <- read_dta("training.dta")
```

## Creat the variable

Notation:

Variable name: student_r

Created from: occup1_r, occup2_r

Label: student_r — yes-is a student, no-not a student, NA-occup1_r and occup2_r are NAs

```{r}
student_data = training %>%
  as_data_frame() %>%
  mutate(student_r = ifelse(occup1_r == 8 | occup2_r == 8, "yes", "no"),
         student_r = ifelse(is.na(student_r) & occup1_r == 20, "no", student_r))
```

There are 260769 observations and 58 variables in original dataset. As we are interested in the `student_r` variables, I choose observations with age between 15-19 and variables related with the `student_r` variables which are ageyrs_r, sex, student_r, area, educate_r, currmarr_r, sexp1yr, SEScat. This subset of dataset is named as `impu_data`.

```{r}
# for sexp1yr > 92, make them as NAs
impu_data = student_data %>% 
  filter(ageyrs_r <= 19) %>% 
  mutate(student_r = as.factor(student_r),
         visit = as.factor(visit),
         sex = ifelse(female==1, "female", "male")) %>% 
  select(ageyrs_r, sex, student_r, area, educate_r, currmarr_r, sexp1yr, SEScat) %>% 
  mutate(sex = as.factor(sex),
         area = as.factor(area),
         educate_r = as.factor(educate_r),
         currmarr_r = as.factor(currmarr_r),
         sexp1yr = ifelse(sexp1yr > 92, NA, sexp1yr),  
         SEScat = as.factor(SEScat))
head(impu_data)
```

```{r}
summary(impu_data)
```

Above are summary of `impu_data`. There are 6208 missing values in `student_r` variables.

# Imputation

# missForest

First, I use `missForest` package to do multiple imputation. This package use random forest algorithm to impute data and will give the optimal imputation result with lowest estimated error.

```{r, message=FALSE}
# need to first change dataset into matrix, then change it into data frame
# how to decide parameters? cross-validation?
impu_data = as.data.frame(as.matrix(impu_data))
set.seed(123)
impu_forest = missForest(impu_data, variablewise = TRUE)
```

Following are summary and head of the imputed data. This data is stored in file "impu_forest.csv".

```{r}
#impu_forest$OOBerror
impu_forest_df = impu_forest$ximp
summary(impu_forest_df)
head(impu_forest_df)
write.csv(impu_forest_df, file = "impu_forest.csv")
```

Percentages of "yes" in different sexes, ages and rounds are shown as following.

```{r}
visit_data = student_data %>% 
  filter(ageyrs_r <= 19)
impu_forest_df$visit = visit_data$visit

table1 = impu_forest_df %>%
  group_by(visit, student_r, sex, ageyrs_r) %>%
  dplyr::summarize(count = n()) %>%
  spread(key = student_r, value = count) %>% 
  mutate(sum = no + yes) %>% 
  mutate(no_prc = round(no/sum, 4)*100, yes_prc = round(yes/sum, 4)*100) %>%
  select(sex, visit, ageyrs_r, no, no_prc, yes, yes_prc, sum) %>% 
  ungroup()

table1 %>%
  select(sex, visit, ageyrs_r, yes_prc) %>% 
  spread(key = ageyrs_r, value = yes_prc) %>% 
  knitr::kable(digits = 3)
```

# Hmisc

I also use the `Hmisc` package to impute data. The `Hmisc` package uses additive semiparametric models to do multiple imputation. Following show the summary and head of first imputation results. This imputed data is stored in file "impu_himsc.csv".

```{r}
set.seed(123)
impu_himsc = aregImpute(~ ageyrs_r + sex + student_r + area + educate_r + currmarr_r + SEScat, data = impu_data)
impu_himsc_l = impute.transcan(impu_himsc, data=impu_data, imputation=1, list.out=TRUE, pr=FALSE, check=FALSE)
impu_himsc_df = as.data.frame(impu_himsc_l)
summary(impu_himsc_df)
head(impu_himsc_df)
write.csv(impu_himsc_df, file = "impu_himsc.csv")
```

Percentages of "yes" in different sexes, ages and rounds are shown as following.

```{r}
impu_himsc_df$visit = visit_data$visit
table2 = impu_himsc_df %>%
  group_by(visit, student_r, sex, ageyrs_r) %>%
  dplyr::summarize(count = n()) %>%
  spread(key = student_r, value = count) %>% 
  mutate(sum = no + yes) %>% 
  mutate(no_prc = round(no/sum, 4)*100, yes_prc = round(yes/sum, 4)*100) %>%
  select(sex, visit, ageyrs_r, no, no_prc, yes, yes_prc, sum) %>% 
  ungroup()

table2 %>%
  select(sex, visit, ageyrs_r, yes_prc) %>% 
  spread(key = ageyrs_r, value = yes_prc) %>% 
  knitr::kable(digits = 3)
```

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

